---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

We are going to load, extract and read the data from the dataset [Activity monitoring data](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip)

```{r}
# This section deals with the downloading the compressed file and
# extracting it contents.

dataFile <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"

download.file(dataFile, destfile = "./activity.zip")

unzip(zipfile = "./activity.zip", exdir = "./activity")

```

The above step is an one time operation, verify if the folder and its contents have been extracted to the appropriate location  once done.

Reading the contents of the uncompressed folder.
```{r}
# Reading the contents of the file
readActivityData <- read.csv("../activity/activity.csv")

summary(readActivityData)

head(readActivityData)

```
Omitting the null values in the data and saving it to a new Variable

```{r}

activityCompleteData <- na.omit(readActivityData)

head(activityCompleteData)

```

## What is mean total number of steps taken per day?

Here we can split this up into 2 parts: 
1) We find the total number of steps. 
2) We find the mean of the total number of steps

```{r}

library(dplyr)

totalSteps <- activityCompleteData %>% select(date, steps) %>% group_by(date) %>% summarize(tsteps= sum(steps))
hist(totalSteps$tsteps, xlab = "Total steps taken daily", ylab = "Count of the number of steps", main="Histogram of Total no of Steps by day", breaks = 20, col = "#FF8509" )

```

Now to calculate the mean of the steps

```{r}

mean(totalSteps$tsteps)


```
Median of the steps
```{r}

median(totalSteps$tsteps)

```

## What is the average daily activity pattern?

Time series plot of the 5-minute interval is as below

```{r}

library(ggplot2)

head(activityCompleteData)

five_min_interval <- activityCompleteData %>% select(interval, steps) %>% na.omit() %>% group_by(interval) %>% summarize(tsteps= mean(steps)) 
  
ggplot(five_min_interval, aes(x=interval, y=tsteps))+ geom_line()

```

```{r}
head(five_min_interval)
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
five_min_interval[which(five_min_interval$tsteps== max(five_min_interval$tsteps)),]
```
## Imputing missing values

Note that there are a number of days/intervals where there are missing values (coded as 𝙽𝙰). The presence of missing days may introduce bias into some calculations or summaries of the data.

1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with 𝙽𝙰s)

```{r}

sapply(X = readActivityData, FUN = function(x) sum(is.na(x)))

```
2. Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc. I will use the mean for that 5 -minute interval to replace all the missing values in the dataset. At the end, I will check if all the NAs have been replaced.

```{r}

replacewithmean <- function(x) replace(x, is.na(x), mean(x, na.rm = TRUE))

meanActivitydata <- readActivityData%>% group_by(interval) %>% mutate(steps= replacewithmean(steps))

head(meanActivitydata)


```

```{r}

sum(is.na(meanActivitydata))
```
3. Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
newActivitydataset <- as.data.frame(meanActivitydata)
head(newActivitydataset)

```

```{r}
summary(newActivitydataset)
```
4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps? We make a histogram of the total number of steps taken each day first by using the new version dataset
```{r}
library(ggplot2)
new_steps <- aggregate(newActivitydataset$steps, by = list(newActivitydataset$date), FUN = sum)
names(new_steps)[names(new_steps) == "x"] <- "Total"
names(new_steps)[names(new_steps) == "Group.1"] <- "Date"
hist2 <- ggplot(data = new_steps, aes(Total)) + 
    geom_histogram(binwidth = 1500, colour = "#FF7864") +
    xlab("Total Number of Steps Taken Each Day") +
    ylab("Count") +
    ggtitle("Histogram of the Total Number of Steps Taken Each Day with New Version Dataset")
print(hist2)

```

## Are there differences in activity patterns between weekdays and weekends?
