---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

We are going to load, extract and read the data from the dataset [Activity monitoring data](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip)

```{r}
# This section deals with the downloading the compressed file and
# extracting it contents.

dataFile <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"

download.file(dataFile, destfile = "./activity.zip")

unzip(zipfile = "./activity.zip", exdir = "./activity")

```

The above step is an one time operation, verify if the folder and its contents have been extracted to the appropriate location  once done.

Reading the contents of the uncompressed folder.
```{r}
# Reading the contents of the file
readActivityData <- read.csv("../activity/activity.csv")

summary(readActivityData)

head(readActivityData)

```
Omitting the null values in the data and saving it to a new Variable

```{r}

activityCompleteData <- na.omit(readActivityData)

head(activityCompleteData)

```

## What is mean total number of steps taken per day?



## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
