---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

We are going to load, extract and read the data from the dataset [Activity monitoring data](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip)

```{r}
# This section deals with the downloading the compressed file and
# extracting it contents.

dataFile <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"

download.file(dataFile, destfile = "./activity.zip")

unzip(zipfile = "./activity.zip", exdir = "./activity")

```

The above step is an one time operation, verify if the folder and its contents have been extracted to the appropriate location  once done.

Reading the contents of the uncompressed folder.
```{r}
# Reading the contents of the file
readActivityData <- read.csv("../activity/activity.csv")

summary(readActivityData)

head(readActivityData)

```
Omitting the null values in the data and saving it to a new Variable

```{r}

activityCompleteData <- na.omit(readActivityData)

head(activityCompleteData)

```

## What is mean total number of steps taken per day?

Here we can split this up into 2 parts: 
1) We find the total number of steps. 
2) We find the mean of the total number of steps

```{r}

library(dplyr)

totalSteps <- activityCompleteData %>% select(date, steps) %>% group_by(date) %>% summarize(tsteps= sum(steps))
hist(totalSteps$tsteps, xlab = "Total steps taken daily", ylab = "Count of the number of steps", main="Histogram of Total no of Steps by day", breaks = 20, col = "#FF8509" )

```

Now to calculate the mean of the steps

```{r}

mean(totalSteps$tsteps)


```
Median of the steps
```{r}

median(totalSteps$tsteps)

```

## What is the average daily activity pattern?

Time series plot of the 5-minute interval is as below

```{r}

library(ggplot2)

head(activityCompleteData)

five_min_interval <- activityCompleteData %>% select(interval, steps) %>% na.omit() %>% group_by(interval) %>% summarize(tsteps= mean(steps)) 
  
ggplot(five_min_interval, aes(x=interval, y=tsteps))+ geom_line()

```

```{r}
head(five_min_interval)
```
## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
